---
title: "RWorksheet_Calzado#4c.Rmd"
author: "Michael Angelo S. Calzado"
date: "2024-11-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset `mpg`
   A data frame with 234 rows and 11 variables:
   - **manufacturer**: manufacturer name
   - **model**: model name
   - **displ**: engine displacement, in litres
   - **year**: year of manufacture
   - **cyl**: number of cylinders
   - **trans**: type of transmission
   - **drv**: the type of drive train (f = front-wheel drive, r = rear wheel drive, 4 = 4wd)
   - **cty**: city miles per gallon
   - **hwy**: highway miles per gallon
   - **fl**: fuel type
   - **class**: "type" of car

• Download and open the mpg file. Upload it to your OWN environment
a. Show your solutions on how to import a csv file into the environment.
```{r}
mpgdata <- read.csv("mpg.csv")
mpgdata
```

b. Which variables from mpg dataset are categorical?

 - manufacturer
 - drv
 - cyl
 - model
 - trans
 - fl
 - class

c. Which are continuous variables?
 - displ
 - year
 - cyt
 - hwy
 


2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
```{r}
library(dplyr)

manufacturercount <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(modelcount = n_distinct(model), .groups = 'drop') %>%
  arrange(desc(modelcount))

mostmodelsmanufacturer <- manufacturercount %>% slice(1)

modelvariationcount <- mpgdata %>%
  group_by(model) %>%
  summarise(variationcount = n(), .groups = 'drop') %>%
  arrange(desc(variationcount))

mostvariationsmodel <- modelvariationcount %>% slice(1)

mostmodelsmanufacturer
mostvariationsmodel
```

a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
uniquemodels <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(uniquemodels = list(unique(model))) %>%
  arrange(manufacturer)

print(uniquemodels)
```
b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
model_counts <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(unique_model_count = n_distinct(model)) %>%
  arrange(desc(unique_model_count))


barplot(model_counts$unique_model_count, 
        names.arg = model_counts$manufacturer, 
        las = 2, 
        col = "lightblue", 
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer", 
        ylab = "Number of Unique Models")


library(ggplot2)

ggplot(model_counts, aes(x = reorder(manufacturer, unique_model_count), y = unique_model_count)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer", 
       x = "Manufacturer", 
       y = "Number of Unique Models") +
  theme_minimal()

```
2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
uniqueModels <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(uniqueModels = list(unique(model))) %>%
  arrange(manufacturer)

print(uniqueModels)
```
b. For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}
modelCounts <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(unique_model_count = n_distinct(model)) %>%
  arrange(desc(unique_model_count))


barplot(model_counts$unique_model_count, 
        names.arg = modelCounts$manufacturer, 
        las = 2, 
        col = "lightblue", 
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer", 
        ylab = "Number of Unique Models")


library(ggplot2)

ggplot(model_counts, aes(x = reorder(manufacturer, unique_model_count), y = unique_model_count)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer", 
       x = "Manufacturer", 
       y = "Number of Unique Models") +
  theme_minimal()

```
3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.
```{r}
library(ggplot2)
library(dplyr)
top_20_mpg <- mpgdata %>%
  arrange(desc(cty)) %>%
  head(20)
ggplot(top_20_mpg, aes(x = model, y = year)) +
  geom_point() +
  labs(title = "Top 20 Models by Year", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}
model_count <- mpgdata %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars))

print(model_count)

```

a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and
colors. Show code and results.
```{r}
top_models <- mpgdata %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars)) %>%
  slice_head(n = 20)

ggplot(top_models, aes(x = reorder(model, number_of_cars), y = number_of_cars, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +  
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Model") +
  theme_minimal() +
  scale_fill_viridis_d() 
```
b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}
library(viridis)
library(viridisLite)
top_model <- mpgdata %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars)) %>%
  slice_head(n = 20)

ggplot(top_model, aes(x = reorder(model, number_of_cars), y = number_of_cars, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  theme_minimal() +
  scale_fill_viridis_d()
```

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.

a. How would you describe its relationship? Show the codes and its result.
```{r}
ggplot(mpgdata, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +  
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  theme_minimal() +
  scale_color_viridis_c() 
```

General Trend: The scatter plot usually shows a positive relationship between the number of cylinders and engine displacement; as the number of cylinders goes up, so does the engine size.

Clustered Data: You’ll likely notice clusters of data points, where cars with fewer cylinders (like 4-cylinder engines) have lower displacement values, while those with more cylinders (like 6 or 8) generally feature larger displacements.

Variability: Although the overall trend is positive, there can be some variability, especially among vehicles with higher cylinder counts, where displacement values may differ significantly.

Outliers: Keep an eye out for outliers that could distort the relationship. For example, a car with a large displacement but only a few cylinders may suggest a performance-focused engine.

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
output?

```{r}
ggplot(mpgdata, aes(x = displ, y = hwy, color = cty)) +
  geom_point(size = 3, alpha = 0.7) + 
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (L)",
       y = "Highway MPG",
       color = "City MPG") +
  theme_minimal() +
  scale_color_viridis_c()
```

Figure 1: Car Model

6. Import the traffic.csv onto your R environment.
```{r}
library(readr)
trafficdata <- read_csv("/cloud/project/Worksheet#4/traffic.csv")
head(trafficdata)
```
a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.
```{r}
numberofobservations <- nrow(trafficdata)
variables <- colnames(trafficdata)
numberofobservations
variables
```
b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}
library(dplyr)
junctions <- subset(trafficdata, Junction == TRUE)
head(junctions)
```
c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
library(ggplot2)
ggplot(trafficdata, aes(x = DateTime, y = Vehicles, color = factor(Junction))) +
  geom_line() +
  labs(title = "Vehicle Counts at Junctions Over Time",
       x = "Date and Time",
       y = "Number of Vehicles",
       color = "Junction") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
7. From alexa_file.xlsx, import it to your environment
a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.
```{r}

library(readxl)
alexafile <- read_excel("alexa_file.xlsx")
dimensions <- dim(alexafile)
rows <- dimensions[1]
columns <- dimensions[2]

rows
columns  

```
b. group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.
```{r}
library(dplyr)
variationtotals <- alexafile %>%
  group_by(variation) %>%
  summarise(total = n())
variationtotals

```
c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.
```{r}
library(ggplot2)

ggplot(variationtotals, aes(x = variation, y = total, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Counts for Each Variation", x = "Variation", y = "Total Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.
```{r}
alexafile$verified_reviews <- as.numeric(alexafile$verified_reviews)
reviewsbydate <- alexafile %>%
  group_by(date) %>%
  summarise(totalverifiedreviews = sum(verified_reviews, na.rm = TRUE))
library(ggplot2)
ggplot(reviewsbydate, aes(x = date, y = totalverifiedreviews)) +
  geom_line(color = "blue") +
  labs(title = "Verified Reviews Over Time", x = "Date", y = "Number of Verified Reviews") +
  theme_minimal()
```
e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a
graph to show its relationship. Show your solution and answer.
```{r}
library(ggplot2)
library(dplyr)

variationratings <- alexafile %>%
  group_by(variation) %>%
  summarise(averagerating = mean(rating, na.rm = TRUE))

ggplot(variationratings, aes(x = variation, y = averagerating, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Rating by Variation", x = "Variation", y = "Average Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```