---
title: "RWorksheet_Calzado#3a.Rmd"
author: "Michael Angelo S. Calzado"
date: "2024-09-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Using Loop Function
for() loop
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value

```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrixB <- matrix(0, nrow = 5, ncol = 5)
for (i in 1:5) {
  for (j in 1:5) {
    matrixB[i, j] <- abs(vectorA[i] - vectorA[j])
  }
}
matrixB

```

2. Print the string "*" using for() function. The output should be the same as shown
in Figure
"*"
"*" "*"
"*" "*" "*"
"*" "*" "*" "*"
"*" "*" "*" "*" "*"
```{r}
n <- 5
for (i in 1:n) {
  rowstring <- paste(rep("*", i))
  print(rowstring)
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
start <- as.numeric(readline(prompt = "Enter the starting number for the Fibonacci sequence: "))
a <- 0
b <- 1
repeat {
  nextfibonacci <- a + b
  if (nextfibonacci > 500) {
    break
  }
  if (nextfibonacci >= start) {
    cat(nextfibonacci, "\n")
  }
  a <- b
  b <- nextfibonacci
}

```
Using Basic Graphics (plot(),barplot(),pie(),hist())
4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result
```{r}
library(readr)
shoes <- read_csv("/cloud/project/Worksheet#4/shoe.csv")
shoes
```
b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}
female_subset <- subset(shoes, Gender == "F")
male_subset <- subset(shoes, Gender == "M")
female_count <- nrow(female_subset)
male_count <- nrow(male_subset)
cat("Female:", female_count, "\n")
cat("Male:", male_count, "\n")
```

c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r}
gender_counts <- table(shoes$Gender)
barplot(gender_counts, 
        main = "Number of Males and Females in Household Data",
        xlab = "Gender",
        ylab = "Count",
        col = c("lightblue", "pink"),
        legend = c("Female", "Male"),
        names.arg = c("Female", "Male"))
```
5. The monthly income of Dela Cruz family was spent on the following:
Food Electricity Savings Miscellaneous
60 10 5 25
a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.
```{r}
expenses <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)
percent_labels <- paste0(names(expenses), " - ", round(expenses / sum(expenses) * 100), "%")
pie(expenses, 
    labels = percent_labels, 
    col = c("lightgreen", "lightblue", "gold", "pink"), 
    main = "Dela Cruz Family Monthly Expenses")

```
6. Use the iris dataset.
data(iris)
a. Check for the structure of the dataset using the str() function. Describe what you
have seen in the output.
```{r}
data(iris)
str(iris)
```
b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r}
colMeans(iris[, 1:4])
```
c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write
the R script and its result.
```{r}
species_counts <- table(iris$Species)
pie(species_counts, 
    main = "Species Distribution in Iris Dataset", 
    col = c("lightblue", "lightgreen", "lightpink"),
    labels = paste(names(species_counts), "\n", species_counts))
legend("topright", 
       legend = names(species_counts), 
       fill = c("lightblue", "lightgreen", "lightpink"))

```
d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.
```{r}
setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")
tail(setosa)
tail(versicolor)
tail(virginica)
```
e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal
width and length, labels for the x and y axis, the pch symbol and colors should be based on
the species.
```{r}
colors <- c("setosa" = "red", "versicolor" = "blue", "virginica" = "green")
symbols <- c("setosa" = 16, "versicolor" = 17, "virginica" = 18)
plot(iris$Sepal.Length, iris$Sepal.Width, 
     col = colors[iris$Species], 
     pch = symbols[iris$Species],
     main = "Iris Dataset", 
     sub = "Sepal Width and Length",
     xlab = "Sepal Length", 
     ylab = "Sepal Width")
legend("topright", legend = c("Setosa", "Versicolor", "Virginica"), 
       col = c("red", "blue", "green"), 
       pch = c(16, 17, 18))

```

Hint: Need to convert to factors the species to store categorical variables.

f. Interpret the result.

Basic Cleaning and Transformation of Objects
---The data structure showed that it is appropriate for both exploratory data analysis and more advanced statistical modeling.
---The mean values provided a quick overview of the characteristics of the iris flowers.
---The pie chart effectively visualized the distribution of species, highlighting the predominance of setosa.
---Subsetting the data allowed for a more granular examination of individual species, which can be crucial for targeted analysis.
---The scatterplot illustrated the interdependence of sepal length and width while visually distinguishing between species, facilitating further exploration into their relationships.


7. Import the alexa-file.xlsx. Check on the variations. Notice that there are ex-
tra whitespaces among black variants (Black Dot, Black Plus, Black Show, Black

Spot). Also on the white variants (White Dot, White Plus, White Show, White
Spot).

Figure 4: Snippet of Alexa Variations
a. Rename the white and black variants by using gsub() function.
Syntax:
RObject$columnName <- gsub("Old Name", "New Name", RObject$columnName)
Write the R scripts and show an example of the output by getting a snippet. To embed
an image into Rmd, use the function below:
knitr::include_graphics("file path")
# knitr::include_graphics("file path")
knitr::include_graphics("D:/RScripts/R_Directory/CS101(lectures_23)/BasicsOfR/RBasics/notes/images/m2.JPG")


b. Get the total number of each variations and save it into another object. Save the
object as variations.RData. Write the R scripts. What is its result?
Hint: Use the dplyr package. Make sure to install it before loading the package.
Syntax for dplyr
RObject %>%
count(RObject$columnName)

Sample Output
c. From the variations.RData, create a barplot(). Complete the details of the chart
which include the title, color, labels of each bar.

d. Create a barplot() for the black and white variations. Plot it in 1 frame, side by
side. Complete the details of the chart.
Example:

6