---
title: "RWorksheet_Calzado#4b.Rmd"
author: "Michael Angelo S. Calzado"
date: "2024-09-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg
A data frame with 234 rows and 11 variables:
#’ \describe{
#’ \item{manufacturer}{manufacturer name}
#’ \item{model}{model name}
#’ \item{displ}{engine displacement, in litres}
#’ \item{year}{year of manufacture}
#’ \item{cyl}{number of cylinders}
#’ \item{trans}{type of transmission}
#’ \item{drv}{the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd}
#’ \item{cty}{city miles per gallon}
#’ \item{hwy}{highway miles per gallon}
#’ \item{fl}{fuel type}
#’ \item{class}{"type" of car}
#’ }
"mpg"
• Download and open the mpg file. Upload it to your OWN environment
a. Show your solutions on how to import a csv file into the environment.
```{r}
mpg_data <- read.csv("mpg.csv")
mpg_data
```

b. Which variables from mpg dataset are categorical?

 - manufacturer
 - drv
 - cyl
 - model
 - trans
 - fl
 - class

c. Which are continuous variables?
 - displ
 - year
 - cyt
 - hwy
 


2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
```{r}
manufacturer_count <- mpg_data %>%
  group_by(manufacturer) %>%
  summarise(model_count = n_distinct(model)) %>%
  arrange(desc(model_count))

most_models_manufacturer <- manufacturer_count[1, ]

model_variation_count <- mpg_data %>%
  group_by(model) %>%
  summarise(variation_count = n()) %>%
  arrange(desc(variation_count))

most_variations_model <- model_variation_count[1, ]

most_models_manufacturer
most_variations_model
```
a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
unique_models <- mpg_data %>%
  group_by(manufacturer) %>%
  summarise(unique_models = list(unique(model))) %>%
  arrange(manufacturer)

print(unique_models)
```
b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
model_counts <- mpg_data %>%
  group_by(manufacturer) %>%
  summarise(unique_model_count = n_distinct(model)) %>%
  arrange(desc(unique_model_count))


barplot(model_counts$unique_model_count, 
        names.arg = model_counts$manufacturer, 
        las = 2, 
        col = "lightblue", 
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer", 
        ylab = "Number of Unique Models")


library(ggplot2)

ggplot(model_counts, aes(x = reorder(manufacturer, unique_model_count), y = unique_model_count)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer", 
       x = "Manufacturer", 
       y = "Number of Unique Models") +
  theme_minimal()

```
2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?

b. For you, is it useful? If not, how could you modify the data to make it more informative?

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.
```{r}
library(ggplot2)
library(dplyr)
top_20_mpg <- mpg_data %>%
  arrange(desc(cty)) %>%
  head(20)
ggplot(top_20_mpg, aes(x = model, y = year)) +
  geom_point() +
  labs(title = "Top 20 Models by Year", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}
model_counts <- mpg %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars))

print(model_counts)

```

a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and
colors. Show code and results.
```{r}
top_models <- mpg_data %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars)) %>%
  slice_head(n = 20)

ggplot(top_models, aes(x = reorder(model, number_of_cars), y = number_of_cars, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +  
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Model") +
  theme_minimal() +
  scale_fill_viridis_d() 
```
b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}
library(viridis)
library(viridisLite)
top_models <- mpg_data %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars)) %>%
  slice_head(n = 20)

ggplot(top_models, aes(x = reorder(model, number_of_cars), y = number_of_cars, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  theme_minimal() +
  scale_fill_viridis_d()
```

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.
```{r}
ggplot(mpg_data, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +  
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  theme_minimal() +
  scale_color_viridis_c() 
```
a. How would you describe its relationship? Show the codes and its result.

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
output?

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point(size = 3, alpha = 0.7) + 
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (L)",
       y = "Highway MPG",
       color = "City MPG") +
  theme_minimal() +
  scale_color_viridis_c()
```

Figure 1: Car Model

6. Import the traffic.csv onto your R environment.
```{r}
library(readr)
trafficdata <- read_csv("/cloud/project/Worksheet#4/traffic.csv")
head(trafficdata)
```
a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.
```{r}
num_observations <- nrow(trafficdata)
variables <- colnames(trafficdata)
num_observations
variables
```
b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}
junctions_data <- subset(traffic_data, Junction == TRUE)
head(junctions_data)
```
c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
library(ggplot2)
ggplot(trafficdata, aes(x = DateTime, y = Vehicles, color = factor(Junction))) +
  geom_line() +
  labs(title = "Vehicle Counts at Junctions Over Time",
       x = "Date and Time",
       y = "Number of Vehicles",
       color = "Junction") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
7. From alexa_file.xlsx, import it to your environment
a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.

b. group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.

c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.

3

d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.

e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a
graph to show its relationship. Show your solution and answer.
=============================== END ==================================